---
title: "Graphs with ggplot2"
output: learnr::tutorial
runtime: shiny_prerendered
description: >
  Quelques géométries classiques, puis la fonction ggpubr::ggarrange
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(ggpubr)
data("nutrimenu", package = "debuter")
couleurs <- c("steelblue", "limegreen", "darkolivegreen", "darkorchid")
```

The `ggplot2` package allows you to easily create graphs with a unified syntax. In this tutorial, we are going to practice with the "nutrimenu" data.

## Basic command

The basic command will look like this

```
ggplot(data, aes(x, y)) + geom_****()
```

Where

* `ggplot` draws an empty canvas
* `data` is a data-frame containing all the information needed to draw the graph
* `aes()` is a function which allows to declare the aesthetic parameters of the graph (the coordinates of points, lines or bars, the way to color them, the labels etc ...)
* `geom_****` is a function that specifies a "drawing layer"

In this tutorial, we will see the following functions:

Function | Result
---------|---------
`geom_bar`, `geom_col` | Bar plot
`geom_point`, `geom_jitter` | Scatterplot
`geom_line` | Line plot
`geom_histogram` | Histograms
`geom_boxplot` | Boxplot
`geom_density` | Density plot
`geom_violin` | Violin plot

While doing the exercises, familiarize yourself with the documentation for ggplot2: https://ggplot2.tidyverse.org/

## The nutrimenu data

The data in the `nutrimenu` table contains the nutritional data of `r nrow nutrimenu)` recipes. The table is shown below.

```{r shownutrimenu}
DT::datatable(nutrimenu)
```

The "Type" columns contains the recipe type.

For more information, go to Nutriwi (https://www.nutriwi.com/).

## Represent a discrete variable

Complete the following command to get a bar chart of recipe types

```{r baton, exercise = TRUE}
ggplot( ) + geom_
```

```{r baton-solution}
ggplot(nutrimenu, aes(Type)) + geom_bar()
```

Change the colors and the theme of the graph by completing the next command.

```{r baton2, exercise = TRUE}
ggplot( , aes(fill = )) + 
   geom_ + 
   theme_
```

```{r baton2-solution}
ggplot(nutrimenu , aes(Type, fill = Type)) + 
   geom_bar() + 
   theme_bw()
```

Use the `cut` function as follows to transform the Proteines (continuous) variable into a qualitative variable.

```{r cutProteines, exercise = TRUE}
cut(nutrimenu$Proteines, c(-1, 6, 14))
```

```{r cutProteines-solution}
# There is nothing more to do: just run the command and observe its result
cut(nutrimenu$Proteines, c(-1, 6, 14))
# If you want, you can use the table function on this result
table(cut(nutrimenu$Proteines, c(-1, 6, 14)))
```

Notice how we can use this trick in a graph!

```{r cutProteinesplot, exercise = TRUE}
ggplot(nutrimenu, aes(cut(Proteines, c(-1, 6, 14)))) +
  geom_bar() + 
  theme_bw()
```

It is not very pretty as a command and a little difficult to read. Sometimes, rather than nesting commands, it's better to create a new variable and use that new variable in the graphics command.

## Represent a continuous variable

Complete the following function to carry out a boxplot of the variable `Energie`.

```{r energiebox, exercise = TRUE}
ggplot(nutrimenu) + geom
```

```{r energiebox-solution}
ggplot(nutrimenu, aes(Energie)) + 
  geom_boxplot() + 
  theme_bw()
```

On the same model, make a violin plot

```{r energieviolon, exercise = TRUE}
ggplot(nutrimenu) + geom
```

```{r energieviolon-solution}
ggplot(nutrimenu, aes(x = Energie, y = 0)) + 
  geom_violin() + 
  theme_bw()
```

Overlap the two and adjust the width and color settings to have a nice graph.

```{r energieboxviolon, exercise = TRUE}
ggplot(nutrimenu) + 
  geom_violin() +
  geom_boxplot()
```

```{r energieboxviolon-hint}
## Use the theme function
### - on the axis.text.y element
### - and on the axis.ticks.y element
### - which you will need to turn into element_blank ()
```

```{r energieboxviolon-solution}
ggplot(nutrimenu, aes(x = Energie, y = 0)) + 
  geom_violin(color = NA, fill = "limegreen") + 
  geom_boxplot(width = 0.1, color = "steelblue", fill = "lightgreen", size = 1) + 
  theme_bw() + 
  labs(x = "", y = "Energie (kCal/100 g)") +    
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

Now make a histogram of the lipid composition.

```{r vitamineChisto, exercise = TRUE}
ggplot(nutrimenu) + 
  geom
```

```{r vitamineChisto-solution}
ggplot(nutrimenu, aes(Lipides)) + 
  geom_histogram(breaks = seq(0, 30, 5), color = "white", fill = "goldenrod") + 
  theme_bw() + 
  labs(x = "Lipides (g/100 g)", y = "Nb. d'occurences")
```

## Represent on the same graph a qualitative variable and a quantitative variable

Run the following command then modify it so that each box is a different color.

```{r sucres, exercise = TRUE}
ggplot(nutrimenu, aes(Type, Glucides)) + geom_boxplot()
```

```{r sucres-solution}
ggplot(nutrimenu, aes(Type, Glucides, color = Type)) + 
  geom_boxplot() + 
  theme_bw() + 
  coord_flip() +
  labs(x = "", y = "Glucides (g/100 g)", color = "")
```

I selected 3 colors for you, and used the following command to store them in a vector called `couleurs`.

```{r, eval = FALSE, echo = TRUE}
couleurs <- c("steelblue", "limegreen", "darkolivegreen")
```

This vector is used below. Run this code and observe the result. Then try other colors of your choice: use color names (see the result of the `colors()` function) or hexadecimal codes.

```{r fibres, exercise = TRUE}
couleurs <- c("steelblue", "limegreen", "darkolivegreen")
ggplot(nutrimenu, aes(Fibres, Type, fill = Type)) + 
  geom_violin() + 
  scale_fill_manual(values = couleurs)
```

## Scatterplot

Complete the following command to represent the energy according to the carbohydrate content of the recipes.

```{r energieVitC, exercise = TRUE}
ggplot() + geom
```

```{r energieGluc-solution}
ggplot(nutrimenu, aes(Glucides, Energie)) +
  geom_point() + 
  theme_bw() + 
  labs(y = "Energie (kCal / 100g)", x = "Glucides (g/100 g)")
```

Color the points, first with the Proteines variable and then with the Type variable. What do you notice?

```{r energieVitCcolor, exercise = TRUE}
ggplot() + geom
```


```{r energieVitCcolor-solution}
ggplot(nutrimenu, aes(Glucides, Energie, color = Proteines)) +
  geom_point() + 
  theme_bw() + 
  labs(y = "Energie (kCal / 100g)", y = "Glucides (g/100 g)")

ggplot(nutrimenu, aes(Glucides, Energie, color = Type)) +
  geom_point() + 
  theme_bw() + 
  labs(y = "Energie (kCal / 100g)", y = "Glucides (g/100 g)")

```

You can change colors defined on a quantitative variable with the function `scale_color_continuous`, or `scale_color_gradient`, or `scale_color_gradientn`, or colors defined on a discrete variable with the `scale_color_manual` or `scale_colour_brewer` function. Test these functions on the previous example.


## Combine graphs

The functionalities of the `ggpubr` package allow, among other things, to combine graphs from `ggplot` objects.

Note: other packages exist to combine `ggplot` objects into complex graphs, like `cowplot`, `gridExtra` or `patchwork`, which is a testament to the success of `ggplot2`.

Run the following commands and see how they allow the creation of a complex graph in multiple panels.

```{r ggarrange, exercise = TRUE}
library(ggpubr)
theme_set(theme_bw())

# boxplots
bxp <- ggplot(nutrimenu, aes(Type, Glucides, color = Type)) + geom_boxplot()
# jitter plots
jp <- ggplot(nutrimenu, aes(x = Type, y = Glucides, color = Type)) + geom_jitter()
# violin plots
vp <- ggplot(nutrimenu, aes(Type, Glucides, color = Type, fill = Type)) + geom_violin() 
# violin plots
dens <- ggplot(nutrimenu, aes(Glucides, fill = Type)) + geom_density() 

# Arrange
ggarrange(bxp, jp, vp, dens, ncol = 2, nrow = 2,  common.legend = TRUE, legend = "bottom") 

```



Here are the key arguments of the `ggarrange` function. Remember that you can access this information and more by calling the function's help file with the `?ggarrange` command.

Argument | Use
---------- | --------
`ncol` | Number of columns in the graph grid.
`nrow` | Number of rows in the graph grid.
`widths` | Vector of the widths of the graphs.
`heights` | Vector of the heights of the graphs.
`legend` | Character string specifying the position of the legend (`"top"`, `"bottom"`, `"left"` or `"right"`). To remove the caption, use `"none"`.
`common.legend` | Logical value. The default is `FALSE`. If `TRUE`, a common unique legend will be created.

