---
title: "Packages in R"
output: learnr::tutorial
runtime: shiny_prerendered
description: >
  Quelques éléments sur Bioconductor et GitHub. Différence entre library et require.
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

## Installing packages

There are several *repositories* from which you can install R packages. Here are a few:

  * The CRAN (for *Comprehensive R Archive Network*),
  * *Bioconductor*,
  * *GitHub*.

There are others, but these three are really the most important.

### CRAN

It is the "original" *repository* which provides (at the time of this tutorial) 17618 packages. The list of these packages is available [here](https://cran.r-project.org/web/packages/available_packages_by_name.html).

The command to install a package from this source is

```
install.packages("package_name")
```

This command will install the `package_name` package in the default library of your session. To find out exactly where this library is located, you can run the command

```
.libPaths()
```

The result depends on a lot of things, including which operating system and which version of R you are using. At home, the result is the following:

```
"/Library/Frameworks/R.framework/Versions/4.0/Resources/library"
```

### Bioconductor

This *repository* is essential for users who wish to do bioinformatics with R. It provides 1974 *software packages*, 398 *experiment data packages*, 968 *annotation packages*, and 28 *workflows* on the date when this tutorial has been written. You can find essential *packages* on Bioconductor like [`limma`] (https://bioconductor.org/packages/release/bioc/html/limma.html) for doing differential analysis of -omics data, or again [`org.Hs.eg.db`] (https://bioconductor.org/packages/release/data/annotation/html/org.Hs.eg.db.html) which contains annotations of the human genome. All of this, and more, is available [at this address] (https://bioconductor.org).

To install a Bioconductor package, you will need to use the function

```
BiocManager::install("package_name")
```

Please note, to use this function, you must first install the `BiocManager` package! (with the command `install.packages("BiocManager")`)


### GitHub

Many R programmers (including me) use a version control system to work, and some use GitHub to host their efforts. Recently, it has been possible to install packages directly from these hosting systems. Be careful, sometimes it's the development version of a package that is on GitHub and not the most stable version!

To install a package from GitHub, you can use the command

```
remotes::install_github("username/package_name")
```

To use this command, you must first install the `remotes` package! (with the command `install.packages("remotes")`)

## Loading packages

There are two functions in R that allow you to load packages: `require` and` library`, which are used in the same way.

```
require(package_name)
require("package_name")
library(package_name)
library("package_name")
```

### If the package exists

When the *package* `package_name` exists, both functions allow the use of functions from that package. The difference between the two is in what the function "returns"

   * `library` returns the *packages* already loaded,
   * `require` returns `TRUE`
  
Exercise: assign to an object called `obj1` and` obj2` the result, respectively, of the command `library(limma)` and of the command `require(limma)`.

```{r retourLimma, exercise = TRUE}

```

```{r retourLimma-solution}
obj1 <- library(limma)
obj1
obj2 <- require(limma)
obj2
```

### If the package does not exist

The most important difference between these two functions appears when we want to load a *package* which does not exist, in this case,

  * `library` generates an error,
  * `require` generates a *warning*, and returns the value `FALSE`.

Exercise: assign to an object called `obj1`, resp. `obj2`, the result of the `library(limma)` command and the `require(limma)` command, respectively.

```{r retourLicorne, exercise = TRUE}

```

```{r retourLicorne-solution}
obj1 <- library(licorne)
obj1
obj2 <- require(licorne)
obj2
```

### Which function to use?

There are other differences between the two functions, cf. the common help page of the two functions accessible with the command `?library` or `?require`.

My advice: use `library` as it will generate an error if you forgot to install a *package*!
