---
title: "Learn to use a tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
description: >
  Plus un entraînement au clavier.
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

In this R programming course, we will occasionally use tutorials that will focus on topics covered in class. These interactive tutorials will allow you to practice with very simple exercises.

## Example


For example, use the following interactive window to run the command `2 + 2`, then click the "Run Code" button.

```{r two-plus-two, exercise=TRUE}

```

For some exercises, I will give you hints, and the solution, which you can access by clicking on the "Hints" and "Solution" button. For example: below, create a vector `x` containing the integers from 1 to 10. Click on the "Solution" button to see the solution.


```{r vec110, exercise=TRUE}

```

```{r vec110-solution}
x <- 1:10
```

Below, declare a vector `x` containing the integers from 1 to 10 ** squared **. Click on the successive clues: the last clue will be the solution! 

```{r vec110bis, exercise=TRUE}

```

```{r vec110bis-hint-1}
1:2^2
```

```{r vec110bis-hint-2}
(1:2)^2
```

```{r vec110bis-solution}
x <- (1:10)^2
```

The following sections contain exercises on R data structures, reading data, as well as keyboard practice to remember how to write the essential symbols for programming in R!


## Keyboard training

Let's start with keyboard training!

Indeed, in R, we often need special characters that are not commonly used. This quiz allows you to practice. Enter the requested character string and click on the "Submit" button.

```{r quiz}
quiz(
  question_text("{}",
    answer("{}", correct = TRUE),
  allow_retry = TRUE,
  trim = TRUE
  ),
  question_text("()",
    answer("()", correct = TRUE),
  allow_retry = TRUE,
  trim = TRUE
  ),
  question_text("[]",
  answer("[]", correct = TRUE),
  allow_retry = TRUE,
  trim = TRUE
  ),
  question_text('x|y',
  answer('x|y', correct = TRUE),
  allow_retry = TRUE,
  trim = TRUE
  ),
  question_text("dat$variable",
  answer("dat$variable", correct = TRUE),
  allow_retry = TRUE,
  trim = TRUE
  ),
  question_text("my_tab[,3:6]",
  answer("my_tab[,3:6]", correct = TRUE),
  allow_retry = TRUE,
  trim = TRUE
  ),
  question_text(htmltools::HTML("x > 1"),
  answer(htmltools::HTML("x > 1"), correct = TRUE),
  answer(htmltools::HTML("x>1"), correct = TRUE),
  allow_retry = TRUE,
  trim = TRUE
  ),
  question_text(htmltools::HTML("x & 1"),
  answer(htmltools::HTML("x & 1"), correct = TRUE),
  answer(htmltools::HTML("x&1"), correct = TRUE),
  allow_retry = TRUE,
  trim = TRUE
  ),
  question_text(htmltools::HTML("x < 1"),
  answer(htmltools::HTML("x < 1"), correct = TRUE),
  answer(htmltools::HTML("x<1"), correct = TRUE),
  allow_retry = TRUE,
  trim = TRUE
  ),
  question_text(htmltools::HTML("lm(y~x)"),
  answer(htmltools::HTML("lm(y~x)"), correct = TRUE),
  answer(htmltools::HTML("lm(y ~ x)"), correct = TRUE),
  allow_retry = TRUE,
  trim = TRUE
  ),  
  question_text(htmltools::HTML("dat %>% select(col)"),
  answer(htmltools::HTML("dat %>% select(col)"), correct = TRUE),
  answer(htmltools::HTML("dat%>%select(col)"), correct = TRUE),
  allow_retry = TRUE,
  trim = TRUE
  )
)
```

